---
title: "R_Day2"
author: "Adebola Agboola"
date: "2024-02-20"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#This is what a chunk looks like
```{r}
plot(1:10)


```

###################################
# 1.2 Get to know the data
###################################
# Read in the data
```{r}
# Read in the two files
dat_1 = read.csv("RNAseq_Data_1.csv")
dat_2 = read.csv("RNAseq_Data_2.csv")

# Look at the head() of each data set
head(dat_1)
head(dat_2)
```

# Challenge Question
# What do you observe when you look at dat_1?
#Upon looking at the data above,it look like data attained from genes of specific chromosomes. The table shows various important parts like the gene length start and end and so on.

# What do you observe when you look at dat_2?
#dat_2 looks at the same gens as the dat_1 however it compares DL instead of the BC in dat_1.


# Let's look at the dimensions and column names of dat_1
```{r}
# This function tells you the number of rows and columns
dim(dat_1)

# The function prints the names of each of the columns
colnames(dat_1)
```

###################################
# 2.3 Try out different tables
##################################
# Summary function
```{r}
# Summarize Sample BC_1_RPKM in dat_1
summary(dat_1$BC_1_fpkm)

# Summarize the Gene Length in dat_1
summary(dat_1$GeneLength)
```

# Table
```{r}
# Count the number of chromosomes are represented in the data
table(dat_1$Chrom)

# Count the number of gene types
table(dat_1$GeneType)
```
# Summarise
```{r}
# Load the appropriate packages that are not built into R
library(dplyr)

# Summarize the gene length by chromosome
summary_data1 <- dat_1 %>%
  group_by(Chrom) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength))

summary_data1


# Summarize the gene length by GeneType
summary_data2 <- dat_1 %>%
  group_by(GeneType) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength))

summary_data2
```
# Summarize the gene length by Chromosome for dat_1
# Calculate the average gene length, total gene length, median gene length, and max gene length
```{r}
stat_data1 <- dat_1 %>%
  group_by(Chrom) %>%
  summarise(mean_geneLength = mean(GeneLength), sum_geneLength = sum(GeneLength), median_genelength = median(GeneLength), max_genelength = max(GeneLength))

stat_data1
```


# Join both dat_1 and dat_2 tables together
```{r}
# Join data using left_join()
dat_joined_left = left_join(dat_1, dat_2, by = c("geneID","GeneName","Chrom","Strand", "Start","End","GeneLength", "GeneType", "GeneDescription"))

# Join data using inner_join()
dat_joined_inner = inner_join(dat_1, dat_2, by = c("geneID","GeneName","Chrom","Strand", "Start","End","GeneLength", "GeneType", "GeneDescription"))

# Look at the column names of the data dat_joined_inner
colnames(dat_joined_inner)
# Reorganize data so GeneType and GeneDescripton is at the end of the table
dat_joined_inner = dat_joined_inner[,c(1:13,16:21,14:15)]

head(dat_joined_inner)
```
# Put it all together
```{r}
# Make a new dataset
dat_new = dat_joined_inner

# Calculate the mean of experiment BC
dat_new$BC_fpkm_average = rowMeans(dat_new[,c(8:10)])

# Calculate the standard deviation of BC using the apply() function
dat_new$BC_fpkm_sd = apply(dat_new[,c(8:10)], 1, sd)

# Calculate the mean of experiment BC_DOX
dat_new$BC_DOX_fpkm_average = rowMeans(dat_new[,c(11:13)])
# Calculate the standard deviation of BC_DOX
dat_new$BC_DOX_fpkm_sd = apply(dat_new[,c(11:13)], 1, sd)
# Calculate the mean of experiment DL
dat_new$DL_fpkm_average = rowMeans(dat_new[,c(14:16)])
# Calculate the standard deviation of DL
dat_new$DL_fpkm_sd = apply(dat_new[,c(14:16)], 1, sd)
# Calculate the mean of experiment DL_DOX
dat_new$DL_DOX_fpkm_average = rowMeans(dat_new[,c(17:19)])
# Calculate the standard deviation of DL_DOX
dat_new$DL_DOX_fpkm_sd = apply(dat_new[,c(17:19)], 1, sd)
head(dat_new)
```

# Now save your data to a csv file and download it to your files
```{r}
write.csv(dat_new, "my_analysis.csv")
```

###################################
# STEP 3 GGPLOT
###################################

# 3.1 GGplot Training challenge
# 3.2 GGplot more in-depth walk-though



#######################################
# 3.2 GGplot more in-depth walk-though
#######################################

```{r}
# Load the tidyverse library
library(tidyverse)
# Load the ggplot2 library
library(ggplot2)

# Create a sample data frame
data <- data.frame(
  Group = rep(c("A", "B", "C", "D"), each = 20),
  Value = rnorm(80)
)

# Create the boxplot
ggplot(data, aes(x = Group, y = Value)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("Value") +
  ggtitle("Boxplot of Values by Group")
```

# Get your data for the Genes_FPKM.csv file into R
```{r}
# Read in the Genes_FPKM.csv file and assign it to the object Genes
Genes = read.csv("Genes_FPKM.csv") 
# What are the number of rows and columns in this dataset?
dim(Genes)
# What type of data is this (matrix or dataframe)?
class(Genes) 
#This is a dataframe.
# Look at the top 6 rows of the dataset
head(Genes)

```
# Challenge question
# What do you think this data is showing?
#This data show the expression of GAPDH and H19 and the experiments are done in triplicates.

# Create GAPDH boxplot
```{r}
# Load the ggplot2 library
library(ggplot2)

# Add color pallete
install.packages('palettetown')
library(palettetown)
# https://github.com/timcdlucas/palettetown

# Make colors!
my_colors <- pokepal(pokemon = 137, spread = 4)


# Create the boxplot
ggplot(Genes, aes(x = Experiment.Group, y = GAPDH_FPKM, fill = Experiment.Group)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("GAPDH_FPKM") +
  scale_fill_manual(values = my_colors) +
  ggtitle("FPKM of GAPDH")

t.test(Genes[7:12,2])
```
# Challenge question
# What do you think each element of the ggplot code does?
# ggplot(Genes, aes(x = Experiment.Group, y = GAPDH_FPKM, fill = Experiment.Group)) +
#This part identifies the dataset that the ggplot will use as well as what the x and y axis are

#  geom_boxplot() +
# This just states what type of boxplot to be used.

#  xlab("Group") +
#Thus is the label for the x axis

#  ylab("GAPDH_FPKM") +
#This is the lable for the y axis

#  scale_fill_manual(values = my_colors) +
#This is assigning the colors to be used in the plot

#  ggtitle("FPKM of GAPDH")
#This is the title of the entire plot.

# Create H19 boxplot
# Be sure to include appropriate comments and code
# Use different colors than for the GAPDH plot
```{r}
my_colors <- pokepal(pokemon = 32, spread = 4)


# Create the boxplot
ggplot(Genes, aes(x = Experiment.Group, y = H19_FPKM, fill = Experiment.Group)) +
  geom_boxplot() +
  xlab("Group") +
  ylab("H19_FPKM") +
  scale_fill_manual(values = my_colors) +
  ggtitle("FPKM of H19")




```

# Challenge question
# Describe what each plot is showing
#Each plot is showing the expression of the specific protien with different experimental groups.

# Compare the two plots to eachother, what is similar and what is different?
#The two plots shows opposite expressions for H19 and GAPDH for BC and DL. It also seems like one is the control.
